\documentclass[12pt]{book}
% setup the page geometry for landscape and use maximum screen real estate
%\usepackage[dvips]{geometry,graphicx,color}
\usepackage[pdftex]{geometry,graphicx,color}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{tensor}
\usepackage{amsfonts}
\usepackage{comment}
\usepackage{srcltx}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{eufrak}
\usepackage{mathtools}
\usepackage{tensor}
\usepackage{bbding}
\usepackage{tensor}
\usepackage{pdfpages}
\usepackage{sectsty}
\usepackage{makecell}
\usepackage{longtable}
\usepackage{float}
\usepackage{listings}
\lstloadlanguages{Python}
%\usepackage[hyphens]{url}
\DeclareGraphicsExtensions{.pdf}
\geometry{headsep=2.0em,hscale=0.80}
\newcommand{\ubh}{\bm{\hat{u}}}
\newcommand{\ebh}{\bm{\hat{e}}}
\newcommand{\ebf}{\bm{e}}
%\definecolor{red}{rgb}{1,0,0}
\setlength{\parindent}{0in}
\usepackage{sectsty}
\sectionfont{\large}
%\titlespacing*{\section}{0pt}{6pt}{3pt}
\include{bookmacros}
\title{\bf\Large An Introduction to Geometric Algebra and Calculus}
\author{\bf Alan Bromborsky\\
\bf Army Research Lab (Retired)\\
\bf abrombo@verizon.net}

\begin{document}
\parskip 10pt

\maketitle
{\bf Introduction}\newline
Geometric algebra is the Clifford algebra of a finite dimensional
vector space over real scalars cast in a form most appropriate for
physics and engineering. This was done by David Hestenes (Arizona
State University) in the 1960's. From this start he developed the
geometric calculus whose fundamental theorem includes the generalized
Stokes theorem, the residue theorem, and new integral theorems
not realized before. Hestenes likes to say he was motivated by the
fact that physicists and engineers did not know how to multiply
vectors.

Researchers at Arizona State and Cambridge have applied these developments 
to classical mechanics,quantum mechanics, general relativity (gauge theory 
of gravity), projective geometry, conformal geometry, etc. 

\tableofcontents

\include{BasicGA}
\include{ExamplesGA}
\include{GeometricCalculusDerivative}
\include{GeometricCalculusIntegration}
\include{GeometricCalculusManifolds}
\include{DifferentialForms}
\include{MultivectorCalculus}
\include{Tensors}
\include{Lagrangian}
\include{LieGroupsAsSpinGroups}
\include{ElectromagneticTheory}
\include{Appendix}

\begin{thebibliography}{99}
\bibitem {Cambridge} Cambridge site: \url{http://www.mrao.cam.ac.uk/~clifford}
\bibitem {AzSt} Arizona State site: \url{http://modelingnts.la.asu.edu}
\bibitem {D&L} Chris Doran and Anthony Lasenby, ``Geometric Algebra for Physicists,'' Cambridge University
Press, 2003.
\bibitem {H&S} David Hestenes and Garret Sobczyk, ``Clifford Algebra to Geometric Calculus,'' Kluwer Academic
Publishers, 1984.
\bibitem {DHS&VA} C. Doran,D. Hestenes, F. Sommen and N. Van Acker, ``Lie Groups as Spin Groups,'' 
	\emph{J. Math. Phys.},\textbf{ 34}(8) August 1993, pp. 3642-3669.
\bibitem {NFCM} David Hestenes, ``New Foundations for Classical Mechanics ($2^{nd}$ edition),'' Kluwer Academic
Publishers, 1999.
\bibitem {Wyss} Walter Wyss,``The Energy-Momentum Tensor in Classical Field Theory,'' \url{http://merlin.fic.uni.lodz.pl/concepts/2005_3_4/2005_3_4_295.pdf}
\end{thebibliography}
\end{document}
