\documentclass[12pt]{book}
% setup the page geometry for landscape and use maximum screen real estate
%\usepackage[dvips]{geometry,graphicx,color}
\usepackage[pdftex]{geometry,graphicx,color}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{tensor}
\usepackage{amsfonts}
\usepackage{comment}
\usepackage{srcltx}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{hyperref}
\geometry{headsep=2.0em,hscale=0.80}
\newcommand{\ubh}{\bm{\hat{u}}}
\newcommand{\ebh}{\bm{\hat{e}}}
\newcommand{\ebf}{\bm{e}}
%\newcommand{\ebh}{\hat{eb}}
%\definecolor{red}{rgb}{1,0,0}
\setlength{\parindent}{0in}
\usepackage{sectsty}
\sectionfont{\large}
%\titlespacing*{\section}{0pt}{6pt}{3pt}
\include{bookmacros}
\title{\bf\Large Manipulations of the Metric}
\author{\bf Alan Bromborsky\\
\bf Army Research Lab (Retired)\\
\bf abrombo@verizon.net}

\begin{document}
\parskip 10pt
\maketitle
Let us consider three different ways of realizing a manifold which I call ``implicit'', ``derived'', and 
``explicit''.

The ``implicit'' manifold is defined by a set of coordinates $\bm{\theta} = \set{\theta_{1},\dots,\theta_{n}}$, 
a set of basis vectors $\set{\eb_{\theta_{1}},\dots,\eb_{\theta_{n}}}$, 
and a metric tensor $\f{g_{ij}}{\bm{\theta}}$ where the
metric tensor is related to the basis vectors by 
\be
	g_{ij} = \eb_{\theta_{i}}\cdot\eb_{\theta_{j}},
\ee
and the $\eb_{\theta_{i}}$ are implicitly a function of the coordinates.  That is to say for an ``implicit'' manifold
the metric tensor defines the dot products of the basis vectors and not vice versa. This way of defining a 
manifold is usually encountered in general relativity.

The ``derived'' manifold is given by a single vector field, $\f{\bm{X}}{\bm{\theta}}$, that is a function of 
the coordinates that defined in some embedding vector space. In this case the basis vectors, 
$\set{\eb_{\theta_{1}},\dots,\eb_{\theta_{n}}}$ and the metric tensor, $\f{g_{ij}}{\bm{\theta}}$, are derived
from $\f{\bm{X}}{\bm{\theta}}$ via
\begin{align}
	\eb_{\theta_{i}} &= \pdiff{\bm{X}}{\theta_{i}} \\
	g_{ij} &= \eb_{\theta_{i}}\cdot\eb_{\theta_{j}},
\end{align}
where the dot product is that of the embedding vector space. In this case the basis vectors and the metric tensor
are derived from $\f{\bm{X}}{\bm{\theta}}$. An example of this would be a two dimensional manifold embedded in
a three dimensional space such as a sphere defined by
$$\f{\bm{X}}{\theta,\phi} = 
\f{\cos}{\theta}\eb_{z}+\f{\sin}{\theta}\paren{\f{\cos}{\phi}\eb_{x}+\f{\sin}{\phi}\eb_{y}}.$$

The ``explicit'' manifold is one in which the basis vectors are given vector fields, 
$\f{\eb_{\theta_{i}}}{\bm{\theta}}$, in an embedding vector space and the metric tensor is calculated as in the
case of the ``derived'' manifold. An example of this would be the spherical coordinate basis vectors
\begin{align*}
	\eb_{r} &= \f{\cos}{\theta}\eb_{z}+\f{\sin}{\theta}\paren{\f{\cos}{\phi}\eb_{x}+\f{\sin}{\phi}\eb_{y}}, \\
	\eb_{\theta} &= \f{\sin}{\theta}\eb_{z}-\f{\cos}{\theta}\paren{\f{\cos}{\phi}\eb_{x}+\f{\sin}{\phi}\eb_{y}}, \\
	\eb_{\phi} &= \f{-\sin}{\phi}\eb_{x}+\f{\cos}{\phi}\eb_{y}.
\end{align*}
In this case the basis vectors can be normalized vector fields and the metric tensor is calculated as in the
``derived'' case.

In both the ``derived'' and ``explicit'' cases the derivatives of the basis vectors can be calculated by 
direct differentiation of the $\f{\eb_{\theta_{i}}}{\bm{\theta}}$'s. However, in the case of the ``implicit'' manifold
we must proceed as follows.

The derivatives of the basis vectors are calculated from the Christ-Awful symbols, $\f{\Gamma_{ij}^{k}}{\bm{\theta}}$,
by (Einstein summation convention used)
\be
	\pdiff{\eb_{\theta_{i}}}{\theta_{j}} = \f{\Gamma_{ij}^{k}}{\bm{\theta}}\eb_{\theta_{k}}.
\ee
Where the $\f{\Gamma_{ij}^{k}}{\bm{\theta}}$ are derived from the $\f{g_{ij}}{\bm{\theta}}$.

Now let us derive all the relevant manifold quatities in terms of normalized basis vectors.

We define a norm for the basis vectors by
\be
	\abs{\eb_{\theta_{i}}} \equiv \sqrt{\abs{g_{ii}}}
\ee
and
\be
	\eb_{\theta_{i}} = \abs{\eb_{\theta_{i}}}\ebh_{\theta_{i}}
\ee
so that $\ebh_{\theta_{i}}\cdot\ebh_{\theta_{i}} = \pm 1$.

Now consider what happens if we use a normalized basis $\set{\ebh_{\theta_{1}},\dots,\ebh_{\theta_{n}}}$.
\begin{align}
	\pdiff{\eb_{\theta_{i}}}{\theta_{j}} &= 
		\pdiff{}{\theta_{j}}\paren{\abs{\eb_{\theta_{i}}}\ebh_{\theta_{i}}} \nonumber \\
		& = \pdiff{\abs{\eb_{\theta_{i}}}}{\theta_{j}}\ebh_{\theta_{i}}
		    +\abs{\eb_{\theta_{i}}}\pdiff{\ebh_{\theta_{i}}}{\theta_{j}} \nonumber \\
	\pdiff{\ebh_{\theta_{i}}}{\theta_{j}} &=
			\bfrac{1}{\abs{\eb_{\theta_{i}}}}\paren{\pdiff{\eb_{\theta_{i}}}{\theta_{j}}
			- \pdiff{\abs{\eb_{\theta_{i}}}}{\theta_{j}}\ebh_{\theta_{i}}},
\end{align}
but
\begin{align}
	\pdiff{\abs{\eb_{\theta_{i}}}^{2}}{\theta_{j}} &= 
		2\abs{\eb_{\theta_{i}}}\pdiff{\abs{\eb_{\theta_{i}}}}{\theta_{j}} \nonumber \\
	\pdiff{\abs{\eb_{\theta_{i}}}}{\theta_{j}} &= \bfrac{1}{2\abs{\eb_{\theta_{i}}}}
	                                              \pdiff{\abs{\eb_{\theta_{i}}}^{2}}{\theta_{j}} \nonumber \\
	                                           &= \bfrac{\abs{\eb_{\theta_{i}}}}{2\abs{g_{ii}}}
	                                              \pdiff{\abs{g_{ii}}}{\theta_{j}}.
\end{align}
So that
\begin{align}
	\pdiff{\ebh_{\theta_{i}}}{\theta_{j}} &= 
		\bfrac{1}{\abs{\eb_{\theta_{i}}}}\pdiff{\eb_{\theta_{i}}}{\theta_{j}}
		- \bfrac{1}{2\abs{g_{ii}}}\pdiff{\abs{g_{ii}}}{\theta_{j}}\ebh_{\theta_{i}} \nonumber \\
	    &= \bfrac{\abs{\eb_{\theta_{k}}}}{\abs{\eb_{\theta_{i}}}}\Gamma_{ij}^{k}\ebh_{\theta_{k}}
	      - \bfrac{1}{2\abs{g_{ii}}}\pdiff{\abs{g_{ii}}}{\theta_{j}}\ebh_{\theta_{i}}.
\end{align}
Note that the derivatives of the normalized basis vectors are not even in the same direction as the derivatives
of the original basis vectors.

My thesis is if the $\eb_{\theta_{i}}$'s are normalized you cannot calculate the 
$\pdiff{\eb_{\theta_{i}}}{\theta_{j}}$'s unless you explicitly know the
 $\f{\eb_{\theta_{i}}}{\bm{\theta}}$ as vector fields in an embedding vector space.  This can be done for
``derived'' and ``explicit'' manifolds, but not for ``implicit'' manifolds. For ``implicit'' manifolds you can 
derive a set of normalized basis vectors and their derivatives in terms of the normalized basis vectors from
a metric for unnormalized basis vectors as shown above.
 
An exception to my thesis could be a case of non-orthogonal basis vectors for then the off diagonal elements of
the metric tensor could encode the necessary information or again maybe not?  

\end{document}
